---
title: "Lab Assignment 1"
author: "Parker DeBruyne - V00837207"
date: "Feb 2, 2024"
output:
  pdf_document: default
---

# Likelihood methods for the Poisson distribution

* Put your name in the author section and fill in the date section above.
* Write R code in the R chunks provided to answer the questions posed.
* Execute each chunk of code to ensure that your code works properly.
* If you cannot compile all of your code without errors before the end of the lab, 
comment out the chunks that are not working.
* Save the Rmd file in your personal directory (refer to Lab 0 instructions if needed)
* knit to pdf (or Word/HTML, but then print those to pdf)
* Upload the pdf file to the Lab Assignment 1 Activity in Brightspace. 

## 1.  First generate 1 observation from the Poisson(lambda=5) distribution and print the value.


```{r data}
set.seed(12345)   #use this seed!

pdat <- rpois(1, lambda = 5)
pdat
```

## 2.  Compute the Log-likelihood for a vector sequence of lambda values from 2 to 7 in steps of .05 .
(Hint:  See help for the R function called *dpois*.)

```{r LogLik}
lambda <- seq(2,7,by=1)

ploglike <- dpois(pdat, lambda, log = TRUE)
ploglike
```


## 3. Plot the Log-Likelihood function values from your answer in 2 versus lambda. Axes must be labelled and the plot must have a title.

```{r plotLog}
head(cbind(lambda, ploglike))


plot(ploglike ~ lambda, ylab='Log-Likelihood', xlab='lambda', type='b', main='Poisson Log-Likelihood for Lab Assignment 1')


```


## 4.  Generate n=10 observations from the same distribution and compute the sample mean of the 10 observations.  
(Hint:  See help for the R function called *mean*.)

```{r mean}
set.seed(12345) #use this seed!
n <- 10
pdat3 <- rpois(n, lambda = 5)
pdat3

sample_mean = mean(pdat3)
sample_mean
```

## 5.  Plot the log-likelihood function given the n=10 observations.  Axes must be labelled and the plot must have a title.

```{r plotLog10}
lambda <- seq(2,7,by=1)

ploglike3 <- 0
for (i in 1:n){
  ploglike3 <- ploglike3 + dpois(pdat3[i], lambda, log=TRUE)
}

head(cbind(lambda, ploglike3))

plot(ploglike3 ~ lambda, ylab='Log-Likelihood', xlab='lambda', type='b', main='Poisson Log-Likelihood for Lab Assignment 1')
```


