---
title: "DC Notes - Useful Functions"
author: "Parker DeBruyne"
date: "01/05/2022"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/admin/Documents/School/STAT 123/Data Camp Notes")
```

n() #size of group

str()
is.*()
as.*()
identical()
sort()

unlist()

rep() #repeat
seq() #sequence
diff() #Returns suitably lagged and iterated differences.

append()
rev() #reverse

abs() #absolute value
sum()
mean()
round()

grep() #search string
grepl() #logical search
sub() #replace
gsub() #replace all
  meta: 
    "x|y" #OR
    "^" #start of string
    "$" #end of string
    ".*" #any character(s) before
    "\\" #escape
    [0-9]+ #match any numbers between, once
    ([0-9]+) #match any and send within function
                 \\1 #recieve

Sys.Date() #today's date
Sys.time() #current time

unclass() #strips format, gives numeric
as.Date("yyyy-mm-dd", format= "%inputmeta-%inputmeta-%inputmeta")
    meta:
      %Y #4-digit year
      %y #2-digit-year
      %m #2-digit month
      %d #2-digit day
      %A #weekday
      %a #abbreviated weekday
      %B #month
      %b #abbreviated month
format(x, format = "%meta-%meta-%meta") 
    meta:
      #see above meta
as.POSIXct(x, format="%meta") #convert string to POSIXct object
  meta:
    %H #hours as a decimal number (00-23) 24hr
    %I #hours as a decimal number (01-12) 12hr
    %M #minutes as a decimal number
    %S #seconds as a decimal number
    %T #shorthand notation for %H:%M:%S
    %p #AM/PM indicator
    full list: ?strptime
    default: %Y-%m-%d %H:%M:%S
          

```{r}   
#Once something is a POSIXct object, R knows how to index it. Use format() to make a substring.

          # Definition of character strings representing times
str1 <- "May 23, '96 hours:23 minutes:01 seconds:45"
str2 <- "2012-3-12 14:23:08"

# Convert the strings to POSIXct objects: time1, time2
time1 <- as.POSIXct(str1, format = "%B %d, '%y hours:%H minutes:%M seconds:%S")
time2 <- as.POSIXct(str2, format = "%Y-%m-%d %T")

# Convert times to formatted strings
format(time1, format = "%M")

format(time2, format="%I:%M %p")
```


#DATA MANIPULATION WITH DPLYR
library(pillar)
library(gapminder)

glimpse() #shows a few values of all the columns
%>%
select( col, -col, col:col, , contains("string"), starts_with("string"), ends_with("string"), last_col(), , )    # ?select_helpers
filter(bool, bool, bool)
arrange(desc())
mutate(new_col = expression)

```{r}
#Aggregate data: to take many observations and summarize them into one.
```
dplyr
%>%
  count(variable, wt = attribute, sort = T/F) #gives the number of observations by a variable. wt the n column will be weighted by a different variable instead of a numeric count. Sort from most common to least.

summarize(new_var = function(var))
arrange(desc(variable)) 
group_by(var, var) #allows for subgroups
ungroup()
filter()
top_n(n_observations, weight_variable) #returns the top value(s) of weight
rename(new_col_name = old_col_name)
