---
title: "STAT 261, Lab 1"
author: "Parker DeBruyne, Arjun Banik"
date: "January 26, 2024"
output:
  pdf_document: default
---

# Likelihood methods for the Binomial distribution

* Put your name in the author section above.
* Execute each chunk of code to ensure that your code works properly.  The TA will review the code with you.
* Sometimes one of your chunks of code will not compile properly, but you must hand
your document in.  In that case, 'Comment' out the R code that is not working properly using # as the first character in your lines of code. Do that now for the chunk of code in Section number 4, below. Put #  at the start of each line of R code. 
* Complete question 6 below. Execute your question 6 code chunk to ensure that it works properly.  If you cannot compile it without errors before the end of the lab, comment out the chunk.
* Save the Rmd file
* Knit the Rmd file to pdf (either directly or by knitting to Word/HTML and printing to pdf)
* You are welcome to submit this pdf into the "Lab 0 (practice submission)" activity if you have not already completed the practice submission, or if you want your instructor to look at your work.

## 1.  First we generate 1 observation from the Binomial(100, theta=.07) distribution


```{r data}
#This is a comment inside of a chunk of R code.

set.seed(54321)   #setting the random number generator seed 
                  #SHOULD always be done before simulating.
# Ask your Lab instructor to explain set.seed()

bdat <- rbinom(1, size=100, prob=.07 )   # see ?rbinom for information
bdat
```

## 2.  We know that the MLE for the probability of success, theta, is the sample proportion.

```{r MLE_theta}
thetahat <- bdat/100
thetahat
```

The MLE of theta is `r thetahat`.  

## 3.  Compute the vector of Log-Likelihoods for a sequence of theta values given the one data point bdat.


```{r Log-Like}
#create a vector of values from 0.02 to 0.14 in steps of 0.005
theta <- seq(0.02, 0.14, by=.005)

#compute the log of binomial probabilities
bloglike <- dbinom(bdat, size=100, prob=theta, log=TRUE)

#combine the theta values and the log probabilities as column vectors
head(cbind(theta, bloglike))
```

## 4. Plot the Log-Likelihood function versus theta.

There are several ways to call the *plot* function in R.  Here are two
ways to plot **Y versus X**.  
1.  plot(x, y)  
2.  plot(y ~ x)    
Where y is plotted on the vertical axis and x is plotted on horizontal axis.

```{r plot}
plot(bloglike ~ theta, ylab='Log-Likelihood', xlab='theta', type='b')
title('Binomial Log-likelihood for Lab 1')
```

## 5.  Generate n=10 observations from the same distribution and plot the log-likelihood function given the n observations.

```{r Sample}
set.seed(54321)
n <- 10
bdat2 <- rbinom(n, size=100, prob=.07 )   # this generates n binomial observations
bdat2
theta <- seq(0.02, 0.14, by=.005)

bloglike2 <- 0
for (i in 1:n){
  bloglike2 <- bloglike2 + dbinom(bdat2[i], size=100, prob=theta, log=TRUE)  
  #computes the sum of the log of binomial probabilities over the observations
} 

head(cbind(theta, bloglike2))

plot(bloglike2 ~ theta, ylab='Log-Likelihood', xlab='theta', type='b')
title(paste('Binomial Log-likelihood for Lab 1', n='', n, sep=''))
```

## 6. Now it is your turn.  In the R code chunk below, redo question 5 for n=100 observations from the same distribution.  How does the log-liklihood function compare with that in question 5?  Write your comments in the section below the R chunk.

```{r YourTurn}
set.seed(54321)
n <- 100
bdat3 <- rbinom(n, size=100, prob=0.07)
bdat3

theta <- seq(0.02, 0.14, by=.005)

bloglike3 <- 0
for (i in 1:n){
  bloglike3 <- bloglike3 + dbinom(bdat3[i], size=100, prob=theta, log=TRUE)
}

head(cbind(theta, bloglike3))

plot(bloglike3 ~ theta, ylab='Log-Likelihood', xlab='theta', type='b')
title(paste('Binomial Log-likelihood for Lab 1', n='', n, sep=''))
```

Comments:  



