---
title: "Lab 2"
author: "Parker DeBruyne - V00837207"
date: "May 12, 2023"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Likelihood methods for the Binomial distribution

* Put your name in the author section above.
* Execute each chunk of code to ensure that your code works properly.  The TA will review the code with you.
* Sometimes one of your chucks of code will not compile properly, but you must hand
your document in.  In that case, 'Comment' out the R code that is not working properly using #
as the first character in your lines of code.
Do that now for the chunk of code in Section number 4, below. Put #  at the start of each line of R 
code. 
* Complete question 6 below.  Execute your question 6 code chunk to ensure that it works properly. If you cannot compile it without errors before the end of the class, comment out the chunk.
* Save this .Rmd file to the (Windows) C: drive  so that there are no knitting issues and then knit the entire document to pdf.
* Copy your pdf and Rmd files to your personal directory (see the Lab 1 instructions if needed).
* Upload the pdf file to the Lab 2 Practice Assignment Activity in the 
Lab section of BrightSpaces before Thursday 11:59pm. 


## 1.  First we generate 1 observation from the Binomial(100, theta=.07) distribution


```{r data}
#This is a comment inside of a chunk of R code.

set.seed(54321)   #setting the random number generator seed 
                  #SHOULD always be done before simulating.
# Ask your Lab instructor to explain set.seed()

bdat <- rbinom(1, size=100, prob=.07 )   # see ?rbinom for information
bdat
```

## 2.  We know that the MLE for the probability of success, theta, is the sample proportion.

```{r MLE_theta}
thetahat <- bdat/100
thetahat
```

The MLE of theta is `r thetahat`.  

## 3.  Compute the vector of Log-Likelihoods for a sequence of theta values given the one data point bdat.


```{r Log-Like}
#create a vector of values from 0.02 to 0.14 in steps of 0.005
theta <- seq(0.02, 0.14, by=.005)

#compute the log of binomial probabilities
bloglike <- dbinom(bdat, size=100, prob=theta, log=TRUE)

#combine the theta values and the log probabilities as column vectors
head(cbind(theta, bloglike))
```

## 4. Plot the Log-Likelihood function versus theta.

There are several ways to call the *plot* function in R.  Here are two
ways to plot **Y versus X**.  
1.  plot(x, y)  
2.  plot(y ~ x)
Where y is plotted on the vertical axis and x is plotted on horizontal axis.

```{r plot1}
#plot(bloglike ~ theta, ylab='Log-Likelihood', xlab='theta', type='b')
#title('Binomial Log-likelihood for Lab 2')

#optional
#abline(v=0.06)
```

```{r plot}
plot(bloglike ~ theta, ylab='Log-Likelihood', xlab='theta', type='b')
title('Binomial Log-likelihood for Lab 2')

#optional
abline(v=0.06) #showing the maximum point with a verticle line
```


## 5.  Generate n=10 observations from the same distribution and plot the log-likelihood function given the n observations.
```{r}
set.seed(54321)
n <- 10
bdat2 <- rbinom(n, size=100, prob=.07 )   # this generates n binomial observations
bdat2
```

```{r}
theta <- seq(0.02, 0.14, by=.005)

bloglike2 <- 0
for (i in 1:n){
  bloglike2 <- bloglike2 + dbinom(bdat2[i], size=100, prob=theta, log=TRUE)  
  #computes the sum of the log of binomial probabilities over the observations
} 

head(cbind(theta, bloglike2))
```


```{r Sample}
plot(bloglike2 ~ theta, ylab='Log-Likelihood', xlab='theta', type='b')
title(paste('Binomial Log-likelihood for Lab 2, n=', n, sep=''))

abline(v=0.06)
```

## 6. Now it is your turn.  In the R code chunk below, redo question 5 for n=100 observations from the same distribution.  How does the log-liklihood function compare with that in question 5?  Write your comments in the section below the R chunk.

```{r YourTurn}
set.seed(54321)
n <- 100
bdat3 <- rbinom(n, size=100, prob=.07)
bdat3
```

```{r}
theta <- seq(0.02, 0.14, by=.005)

bloglike3 <- 0
for (i in 1:n){
  bloglike3 <- bloglike3 + dbinom(bdat3[i], size=100, prob=theta, log=TRUE)
  #computes the sum of the log of binomial probabiliteis over the observations
}

head(cbind(theta, bloglike3))
```

```{r}
plot(bloglike3 ~ theta, ylab='Log-Likelihood', xlab='theta', type='b')
title(paste('Binomial Log-Likelihood for Lab 2, n=', n, sep=''))

abline(v=0.06)
```

Comments:  

The peak has shifted from 0.06
When you are selecting a size more than 40, your distribution is normal; because we are taking more estimates making it more percise.


