---
title: "Lab Assignment 1"
author: "Fill in your Name"
date: "Fill in the date"
output:
  pdf_document: default
  html_document: default
---

# Likelihood methods for the Poisson distribution

* Put your name in the author section above.
* Write R code in the R chunks provided to answer the questions posed.
* Execute each chunk of code to ensure that your code works properly.
* If you cannot compile all of your code without errors, comment out the chunks that are not working.
* Save this .Rmd file to the (Windows) C: drive so that there are no knitting issues and then knit the entire document to pdf.
* Copy your pdf and Rmd files to your personal directory (see the Lab 1 instructions if needed).
* Upload the pdf file to the Lab Assignment 1 in BrightSpace. 

## 1.  First generate 1 observation from the Poisson(lambda=5) distribution and print the value. [1 mark]


```{r data}
set.seed(12345)   #use this seed!


```

## 2.  Compute the Log-likelihood for a vector sequence of lambda values from 2 to 7 in steps of .05. 
(Hint:  See help for the R function called *dpois*.)

```{r LogLik}

```


## 3. Plot the Log-Likelihood function values from your answer in 2 versus lambda. Axes must be labelled and the plot must have a title. [4 marks]

```{r plotLog}

```


## 4.  Generate n=10 observations from the same distribution and compute and display the sample mean of the 10 observations.  (This is the MLE of lambda.) [1 mark]
(Hint:  See help for the R function called *mean*.)

```{r mean10}
set.seed(12345) #use this seed!

```

## 5.  Plot the log-likelihood function given the n=10 observations.  Axes must be labelled and the plot must have a title. [4 marks]

```{r plotLog10}

```

## 6.  Generate n=100 observations from the same distribution and compute and display the sample mean of the 100 observations.  (This is the MLE of lambda.) [1 mark]


```{r mean100}
set.seed(12345) #use this seed!

```

## 7.  Plot the log-likelihood function given the n=100 observations.  Axes must be labelled and the plot must have a title. [4 marks]

```{r plotLog100}

```

