---
title: "Midterm 2"
author: "Parker DeBruyne - V00837207"
date: "21/03/2022"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
---

```{r setup, include=FALSE}
rm(list=ls())
setwd("/Users/admin/Documents/School/STAT 123/Midterms/midterm2")
knitr::opts_chunk$set(echo = TRUE)
```

Question 1: Use the built-in data set HairEyeColor to answer this question.

(a) Create a single table called hair_eye_totals which summarizes the total number of statistics students with each combination of hair and eye colour. 

Note: The built-in data set consists of two tables with this information (one for women and one for men). The answer to part (a) is a single table combining the information from these two tables. 

```{r 1a}
head(HairEyeColor)

HEC = HairEyeColor
#want total students of each combination (either gender)
hair_eye_totals = HEC[,,1]+HEC[,,2]

```

(b) Print out the hair_eye_totals table. 

```{r 1b}
hair_eye_totals
```

(c) Create a grouped bar plot which displays the information from the hair_eye_totals table. Your plot should include the following:
- a main title
- titles for the x-axis and y-axis
- colours to help differentiate the bars 
- a legend to identify what each colour represents

```{r 1c}
barplot(hair_eye_totals,legend=rownames(hair_eye_totals),
        main="Hair and Eye Colour Combinations of Students", 
        col = c("black","brown","red","yellow"),
        xlab ="Hair Colour and Eye Colour", ylab="Frequency",beside=TRUE)
```

(d) Create and print out a vector called percent_eye which contains the percent of statistics students with each eye colour (rounded to 2 decimal places). Show any additional code needed to create this vector. 

```{r }
totals = colSums(hair_eye_totals)
percent_EC = round((totals/sum(totals))*100,2)
percent_EC
```

(e) Create a pie chart displaying the information in the percent_eye vector. Your graph should include:
- a main title
- labels for each wedge displaying the eye colour
- a different colour for each eye colour 
- the percentages displaying next to each wedge. 

```{r }
EC_labels = paste(names(percent_EC),percent_EC," %")
pie(percent_EC,labels=EC_labels,main="Eye Colours of Students ",
    col=c("brown","green","orange","blue"))
```



```{r question 1}

# (a) Answer below:

# (b) Answer below:

# (c) Answer below:

# (d) Answer below:

# (e) Answer below:

```

Question 2: Use the built-in data set USArrests to answer this question. 

(a) What is the variable Murder being measured in the data set? 
```{r }
?USArrests

#Ans: Number of Murder arrests (per 100,000)
```

(b) What type of variable is this?
```{r }
str(USArrests$Murder)

#Ans: Numeric
```

(c) What is the most appropriate type of graph to visualize the distribution of this variable? 
```{r }
#Ans: Histogram
```

(d) Graph the distribution of the variable (using the type of graph that you identified in part (c)). Your graph should include:
- a main title.
- x-axis title.  
- scales on the x and y-axis which fully extend from at least the min value to at least the max value. 
```{r }
hist(USArrests$Murder, main="Frequency of Values", 
     xlab="Number of Murder arrests (per 100,000)")
```

(e) Describe the shape of the distribution (that is, symmetric, left-skewed, right-skewed). 
```{r }
#Ans: The distribution is not symmetric and is right skewed.
```

(f) What is an appropriate statistic to measure the center of the distribution? Why?
```{r }
# Median, because we do not have a normal distribution.
```

(g) Compute the observed value of this statistic. 
```{r }
median(USArrests$Murder)
```

(h) What is an appropriate statistic to measure the spread of the distribution? Why?
```{r }
# Ans: The 1st and 3rd quantiles since we are using the median 
#      as the center of our distribution.
```

(i) Compute the observed value of this statistic. 
```{r }
quantile(USArrests$Murder,c(0.25,0.75))
```

```{r Question 2}

# (a) Answer below:

# (b) Answer below:

# (c) Answer below:

# (d) Answer below:

# (e) Answer below:

# (f) Answer below:

# (g) Answer below:

# (h) Answer below:

# (i) Answer below:


```


Question 3: Suppose you take a random sample of size 100 of a normally distributed variable Z. The sample mean is 126 and the sample standard deviation is 18. 

(a) Between what range of values should approximately 70% of the observations lie? 
```{r }
#68-95-99.7
mu = 126
sig = 18

diff = (100-70)/2
first = diff
last = 100-diff
qnorm(first/100, mean = mu, sd = sig)
qnorm(last/100, mean = mu, sd = sig)
```

(b) Between what range of values should approximately 80% of the observations lie? 
```{r }
mu = 126
sig = 18

diff = (100-88)/2
first = diff
last = 100-diff
qnorm(first/100, mean = mu, sd = sig)
qnorm(last/100, mean = mu, sd = sig)
```

(c) What is the estimated standard error for the sample mean?
```{r }
n=100
std_err = sig/sqrt(n)
std_err
```

(d) What is the critical value for an 86% confidence interval for the mean? 
```{r }
diff_86 = (100-86)/2
interval_86 = c(diff_86, 100-diff_86)

crit_val_86 = qnorm(interval_86/100, mean=0, sd=1)
print(paste("Critical Value estimate for an 86% confidence interval =", round(crit_val_86[2], 2)))
```

(e) Determine an 86% confidence interval for the mean.
```{r }
confidence_int_86 = c(mu - crit_val_86[2]*std_err, mu + crit_val_86[2]*std_err)

print(paste("Confidence interval for 86% = (",
            round(confidence_int_86[1], 2), 
            ",", 
            round(confidence_int_86[2], 2), 
            ")"))


```


```{r Question 3}

# (a) Answer below:

# (b) Answer below:

# (c) Answer below:

# (d) Answer below:

# (e) Answer below:

```



Question 4: Consider the gapminder data set that we worked with in class. We will need this data set to answer this question. 

(a) Either load the data set into R by typing in library(gapminder) or download the gapminder.csv file from Brightspace and read the data into R, saving it as gapminder. 
```{r }
library(gapminder)
```

(b) Suppose you are looking to explore the relationship between the population and Life Expectancy. What type of graph should you use to visualize this relationship? 
```{r }
#Ans: Line Graph as it is a time series
```

(c) Create a graph which visualizes the relationship between these two variables. Put Life Expectancy is on the x-axis. This graph does not need any titles. 
```{r }
plot(gapminder$lifeExp,gapminder$pop,
     xlab = "Life Expectancy",
     ylab = "Population")
```

(d) What is wrong with the graph? 
```{r }
# Ans: It is clustered near the bottom, with a positive relationship between outliers. This implies there is another relationship to consider.
```

(e) Create a vector which contains the populations recorded for Italy in the data set. Call this vector italy_pop. 
```{r }
# vitaly_index = which(gapminder$Country == "Italy")
# italy_population = gapminder$Country[pop]
```

(f) Create a vector which contains the Life Expectancy for Italy in the data set. Call this vector italy_lifexp. 
```{r }

```

(g) Create a graph which visualizes the relationship between the population size (on y-axis) and Life Expectancy (on x-axis) 
for Italy. Your graph should include:
- a main title.
- a title for both the x-axis and the y-axis
- the scale should not be in scientific notation. 
```{r }

```

(h) Describe the direction and form of the relationship. 
```{r }

```

```{r Question 4}

# (a) Answer below:


# (b) Answer below:


# (c) Answer below:


# (d) Answer below:


# (e) Answer below:


# (f) Answer below:


# (g) Answer below:


# (h) Answer below:



```


Question 5: We will again use the data from the built-in data vector USArrests$Murder. 

(a) Create a variable n which equals the sample size for the variable. 
```{r }

```

(b) Bootstrap 10000 sample means and save the bootstrapped means to a vector called mean_Murder. 
```{r }

```

(c) Plot the sampling distribution of the sample mean (with probability = TRUE) and plot an estimated density curve on the same graph. Your plot should include the following:
- a main title
- a title for the x-axis
- a density curve which is a different colour than your plot. 
```{r }

```

(d) What kind of distribution does it look like? Was this what you expected? Explain. 
```{r }

```

(e) Bootstrap 10000 sample 80th percentiles and save the bootstrapped 80th percentiles to a vector called percentile80_Murder. 
```{r }

```

(f) Plot the sampling distribution of the sample 80th percentile. Your plot should include the following: 
- a main title
- a title for the x-axis
```{r }

```

(e) Compute a 96% confidence interval for the 80th percentile. 
```{r }

```

```{r question 5}

# (a) Answer below:

# (b) Answer below:

# (c) Answer below:

# (d) Answer below:

# (e) Answer below:

```




