---
title: "Midterm 2"
author: "Parker DeBruyne - V00837207"
date: "21/03/2022"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
---

```{r setup, include=FALSE}
rm(list=ls())
setwd("/Users/admin/Documents/School/STAT 123/Midterms/midterm2")
knitr::opts_chunk$set(echo = TRUE)
```

Question 1: Use the built-in data set HairEyeColor to answer this question.

(a) Create a single table called hair_eye_totals which summarizes the total number of statistics students with each combination of hair and eye colour. 

Note: The built-in data set consists of two tables with this information (one for women and one for men). The answer to part (a) is a single table combining the information from these two tables. 

```{r 1a}
hair_eyes_totals = HairEyeColor[,,1] + HairEyeColor[,,2]

```

(b) Print out the hair_eye_totals table. 

```{r 1b}
hair_eyes_totals
```

(c) Create a grouped bar plot which displays the information from the hair_eye_totals table. Your plot should include the following:
- a main title
- titles for the x-axis and y-axis
- colours to help differentiate the bars 
- a legend to identify what each colour represents

```{r 1c}
barplot(hair_eyes_totals,
        legend=rownames(hair_eyes_totals),
        main="Hair and Eye Colour Combinations of Students",
        col= c("black", "brown", "red", "yellow"),
        xlab ="Hair Colour and Eye Colour", 
        ylab="Frequency",
        beside=TRUE)
```

(d) Create and print out a vector called percent_eye which contains the percent of statistics students with each eye colour (rounded to 2 decimal places). Show any additional code needed to create this vector. 

```{r }
totals = colSums(hair_eyes_totals)
percent_EC = round(((totals/sum(totals))*100),2)
percent_EC
```

(e) Create a pie chart displaying the information in the percent_eye vector. Your graph should include:
- a main title
- labels for each wedge displaying the eye colour
- a different colour for each eye colour 
- the percentages displaying next to each wedge. 

```{r }
EC_labels = paste(names(percent_EC), percent_EC, "%")
  
pie(percent_EC, 
    main="Student Eye Colours",
    labels=EC_labels,
    col=c("Brown","Blue","Tan","Green"))
```

Question 2: Use the built-in data set USArrests to answer this question. 

(a) What is the variable Murder being measured in the data set? 
```{r }
#Number of murder arrests per 100,000 residents for each 50 states.
```

(b) What type of variable is this?
```{r }
# Population Proportion: numeric.
```

(c) What is the most appropriate type of graph to visualize the distribution of this variable? 
```{r }
#Histogram
```

(d) Graph the distribution of the variable (using the type of graph that you identified in part (c)). Your graph should include:
- a main title.
- x-axis title.  
- scales on the x and y-axis which fully extend from at least the min value to at least the max value. 
```{r }
hist(USArrests$Murder, 
     main="Distribution of Murder Arrests",
     xlab="Arrests per 100,000 residents",
     ylim=c(0,20))
```

(e) Describe the shape of the distribution (that is, symmetric, left-skewed, right-skewed). 
```{r }
#Right-skewed
```

(f) What is an appropriate statistic to measure the center of the distribution? Why?
```{r }
#Median: because the distribution is not normal.
```

(g) Compute the observed value of this statistic. 
```{r }
median_USAmurder = median(USArrests$Murder)
median_USAmurder
```

(h) What is an appropriate statistic to measure the spread of the distribution? Why?
```{r }
#Quartiles Q1 and Q3 since we are using the median as our centre.
```

(i) Compute the observed value of this statistic. 
```{r }
q1 = quantile(USArrests$Murder, 0.25)
q3 = quantile(USArrests$Murder, 0.75)

c(q1, median_USAmurder, q3)
```



Question 3: Suppose you take a random sample of size 100 of a normally distributed variable Z. The sample mean is 126 and the sample standard deviation is 18. 

(a) Between what range of values should approximately 70% of the observations lie? 
```{r }
# (100-70)/2 = 30/2 = 15
qnorm(c(0.15,0.85), mean=126, sd=18)

```

(b) Between what range of values should approximately 80% of the observations lie? 
```{r }
#(100-80)/2 = 20/2 = 10
qnorm(c(0.1,0.9), mean=126, sd=18)
```

(c) What is the estimated standard error for the sample mean?
```{r }
ese_mean = 18/sqrt(100)
ese_mean
```

(d) What is the critical value for an 86% confidence interval for the mean? 
```{r }
#(100-86)/2 = 14/2 = 7
cv = qnorm(0.93)
cv
```

(e) Determine an 86% confidence interval for the mean.
```{r }
ci = c(126 - ese_mean*cv, 126 + ese_mean*cv)
ci
```



Question 4: Consider the gapminder data set that we worked with in class. We will need this data set to answer this question. 

(a) Either load the data set into R by typing in library(gapminder) or download the gapminder.csv file from Brightspace and read the data into R, saving it as gapminder. 
```{r }
library(gapminder)
```

(b) Suppose you are looking to explore the relationship between the population and Life Expectancy. What type of graph should you use to visualize this relationship? 
```{r }
#Scatterplot
```

(c) Create a graph which visualizes the relationship between these two variables. Put Life Expectancy is on the x-axis. This graph does not need any titles. 
```{r }
plot(gapminder$lifeExp,gapminder$pop)
```

(d) What is wrong with the graph? 
```{r }
#The plot is crowded and hard to read, there is a lot of duplication due to multiple country rows.
```

(e) Create a vector which contains the populations recorded for Italy in the data set. Call this vector italy_pop. 
```{r }
italy_index = which(gapminder$country == "Italy")
italy_pop = gapminder$pop[italy_index]
italy_pop
```

(f) Create a vector which contains the Life Expectancy for Italy in the data set. Call this vector italy_lifexp. 
```{r }
italy_lifexp = gapminder$lifeExp[italy_index]
italy_lifexp
```

(g) Create a graph which visualizes the relationship between the population size (on y-axis) and Life Expectancy (on x-axis) 
for Italy. Your graph should include:
- a main title.
- a title for both the x-axis and the y-axis
- the scale should not be in scientific notation. 
```{r }
options(scipen=999)
plot(italy_lifexp,italy_pop,
     main="Populaiton vs. Life Expectancy in Italy",
     xlab="Life Expectancy",
     ylab="Population")
```

(h) Describe the direction and form of the relationship. 
```{r }
#This is a strong, positive, linear relationship.
```


Question 5: We will again use the data from the built-in data vector USArrests$Murder. 

(a) Create a variable n which equals the sample size for the variable. 
```{r }
n = length(USArrests$Murder)
n
```

(b) Bootstrap 10000 sample means and save the bootstrapped means to a vector called mean_Murder. 
```{r }
mean_Murder = numeric()
for (i in 1:10000){
  mean_Murder[i] = mean(sample(USArrests$Murder, n, replace=TRUE))
}

bs_mean_Murder = mean(mean_Murder)
```

(c) Plot the sampling distribution of the sample mean (with probability = TRUE) and plot an estimated density curve on the same graph. Your plot should include the following:
- a main title
- a title for the x-axis
- a density curve which is a different colour than your plot. 
```{r }
hist(mean_Murder,
     main="Sampling Distribution of Murder Mean",
     xlab="Bootstrapped Murder Mean",
     prob= TRUE)
lines(density(mean_Murder), col="red")
```

(d) What kind of distribution does it look like? Was this what you expected? Explain. 
```{r }
# The distribution appears to be normal, which is congruent with Central Limit Theorem
```

(e) Bootstrap 10000 sample 80th percentiles and save the bootstrapped 80th percentiles to a vector called percentile80_Murder. 
```{r }
percentile80_Murder = numeric()
for (i in 1:10000){
  percentile80_Murder[i] = quantile(sample(USArrests$Murder,n,replace=TRUE), 0.8)
}

mean(percentile80_Murder)
```

(f) Plot the sampling distribution of the sample 80th percentile. Your plot should include the following: 
- a main title
- a title for the x-axis
```{r }
hist(percentile80_Murder,
     main="Sampling Distribution of 80th percentile",
     xlab="Bootstrapped 80th percentile",
     prob=TRUE)
lines(density(percentile80_Murder), col="red")
```

(g) Compute a 96% confidence interval for the 80th percentile. 
```{r }
#(100-96)/2 = 4/2 = 2
ci = quantile(percentile80_Murder,c(0.02,0.98))
ci
```




